---
title: "Exploratory Analysis of the Data"
author: "Brian Davies"
date: "16 March 2015"
output: html_document
---

# Basic Setup

```{r basic_setup}
set.seed(773230)
source( "~/SwiftKey/2_ExploratoryAnalysis.R")
library(hash)
```

# Loading Sample Dataset

```{r load_data, cache = TRUE }
load( 'sampleListClean.Rdata' )
```

# Word Frequency

```{r unigram_frequencies, cache = TRUE}
wordVec <- unlist( sampleListClean )
wordFreq <- table( wordVec )
wordFreq <- wordFreq[ order( wordFreq, decreasing = TRUE )]
head( wordFreq, 20 )
wordRunTotal <- cumsum( wordFreq )
head( wordRunTotal, 20 )
hapaxVec <- names( wordFreq[ wordFreq == 1 ] )
nonHapaxVec <- names( wordFreq[ wordFreq > 1 ] )
```

# Hapax flagging

A __hapax legomenon__ is a word that appears only once in a given corpus.

```{r flag_hapax, cache = TRUE }
sampleListHapax <- sapply( sampleListClean,
                           function(x){ ReplaceHapaxInv( x, nonHapaxVec )} )
save( sampleListHapax, file = "sampleListHapax.Rdata" )
```

# Bigrams and Trigrams

Here we count the sentences as being bracketed with BEGIN and END. How to
build the frequency table?

Answer: I go through the source text one bigram or trigram
at a time, counting the frequency of each. This takes AGES and I probably
need to do some tuning, but Ihere's basically what I ran.

__TODO__: Investigate whether __table()__ does it faster/better. i.e. just enumerate
a great big list of ngrams and then count 'em.

```{r get_ngram_lists, eval = FALSE }
wordListClean <- unlist( sampleListClean )
unigramList <- table( wordListClean )
unigramList <- unigramList[ order( unigramList, decreasing = TRUE)]
ngrams <- names( unigramList )
freqs <- unname( unigramList )
rm( unigramList )
unigramFrame <- data.frame( token1 = factor( ngrams ), freq = freqs )
save( unigramFrame, file = 'unigramFrame.Rdata' )
bigramFrame <- NgramFrame( sampleListHapax, n=2 )
save( bigramFrame, file="bigramFrame.Rdata")
trigramFrame <- NgramFrame( sampleListHapax, n=3 )
save( trigramFrame, file="trigramFrame.Rdata")
fourgramFrame <- NgramFrame( sampleListHapax, n=4 )
save( fourgramFrame, file="fourgramFrame.Rdata")
fivegramFrame <- NgramFrame( sampleListHapax, n=5 )
save( fivegramFrame, file="fivegramFrame.Rdata")
trigramPrev <- RearrangeNgramFrame( trigramFrame )
fourgramPrev <- RearrangeNgramFrame( fourgramFrame )
fivegramPrev <- RearrangeNgramFrame( fivegramFrame )
bigramPrev <- bigramFrame
colnames( bigramPrev ) <- colnames( trigramPrev )
unigramPrev <- unigramFrame
colnames( unigramPrev ) <- c( 'nextTok', 'freq' )
save( unigramPrev, file = "unigramPrev.Rdata" )
save( bigramPrev, file = "bigramPrev.Rdata" )
save( trigramPrev, file = "trigramPrev.Rdata" )
save( fourgramPrev, file = "fourgramPrev.Rdata" )
save( fivegramPrev, file = "fivegramPrev.Rdata" )
```

```{r reload_ngram_lists, cache = TRUE}
load( "uniGramList.Rdata" )
load( "biGramList.Rdata" )
load( "triGramList.Rdata" )
```

# Frequency and meta-frequency Analysis

The lists of unigrams, bigrams and trigrams are already labelled with frequency.
Want to sort them all into order of decreasing frequency.

Easy to do for unigrams:

```{r sort_unigrams_descending, cache = TRUE}
unigramsSorted <- uniGramList[ order( uniGramList, decreasing = TRUE )]
head( unigramsSorted, 20 )
```

Needs some coding for bigrams and trigrams because they're hashes.

```{r analyze_bigrams, cache = TRUE}
biGramFrame <- data.frame(
  bigram = keys( biGramList ),
  frequency = values( biGramList )
  )
biGramFrameOrder <- order( biGramFrame$frequency, decreasing = TRUE )
biGramFrame$bigram <- biGramFrame$bigram[biGramFrameOrder]
biGramFrame$frequency <- biGramFrame$frequency[biGramFrameOrder]
head( biGramFrame, 20 )
save( biGramFrame, file = "biGramFrame.Rdata" )
```
